Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-10-20 draft.tex
[draft] Some comments and notes

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[svgnames]{xcolor}
\usepackage{hepparticles}
\usepackage{hepnicenames}
\usepackage{hepunits}
\usepackage{graphicx}
\usepackage{epstopdf}

@@

Integration over the non-angular variables yields
\begin{equation}
    \langle S_i\rangle
    \equiv \int \rmdx{^M \nu} S_i(\nuvec)
    = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N \theta} P(\nuvec,\thvec) \dual{f}_i(\thvec) \right].
\end{equation}
The remainder of this section describes the method of moments, in which we replace the
analytical integration \refeq{det-Pi-analytical} by Monte Carlo (MC) estimators.\\

The central tenet of MC integration is the fact that the expectation value $E_P[g]$ of some function
$g(x)$ under the probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int \rmdx{x} P(x) g(x),
\end{equation}
can be approximated \cite{MCSM-2004} by the consistent and unbiased
estimator $\est{E_P[g]}$~\cite[sec. 8.2]{James:2006zz}
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1}^K g(x^{(k)}) \,,\,    x^{(k)} \sim P
\end{equation}
due to the strong law of large numbers for $K \to \infty$, assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$ are distributed as
$P$.
Throughout this letter we denote all MC estimators with a wide hat.\\

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain observables integrated over certain
bins of $\nuvec$. We define
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & \equiv \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} S_i(\nuvec)\\
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta} P(\nuvec,\thvec) \dual{f}_i(\thvec)  \right]\\
    & = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta}  P(\nuvec,\thvec) \dual{f}_i(\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \right],
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
For notational simplicity, let us forget about the $\nuvec$
integration and consider only $S_i$. In the limit $K \to \infty$, the
central limit theorem (CLT) implies that the random vector
\begin{equation}
  \label{eq:angular-obs-vec}
  \vecest{S} \equiv (S_1, \dots, S_i\,,
  \dots)
\end{equation}
follows a multivariate Gaussian distribution $\mathcal{N}( \vec{S},
\widehat{\Sigma})$ centered on the true value $\vec{S}$ with a
covariance estimated as
\begin{equation}
    \cov[S_i,S_j] \to \est{\cov}[{S}_i, {S}_j]
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[\dual{f}_i(x^{(k)}) - \est{ S_i}\big]\,\big[\dual{f}_j(x^{(k)}) - \widehat{ S_j}\big]\,.
\end{equation}
In our physics applications, the parameter space is compact and each
$\dual{f}_i$ is bounded. Hence the requisites for the most basic
version of the CLT to hold --- finite mean and covariance of
$\dual{f}_i$ --- are automatically satisfied.  We would like to
emphasize that the sample covariance should \fred{``Should'' as in it
  does in practice? Reference toy studies below} rapidly converge
towards the true covariance matrix.  We propose to declare convergence
\fred{I'd prefer: check that skewness and kurtosis are sufficiently
  small} once estimators for skewness and kurtosis are sufficiently
small,

@@

  For a likelihood fit, the constraint needs to be part of the
  analysis performed by the experimental collaboration, and the
  resulting likelihood as a function of $\vec{S}$ is distinctly not
  Gaussian.  For experimentalists, communicating such a result has
  proved to be challenging due to technical reasons (data formats,
  size etc.) and the fear that theorists misinterprete the many
  nuisance parameters (from control channels, detector calibrations
  etc.) that are also part of the likelihood. This leads to the
  undesirable situation that that only the mode and standard errors
  are reported, and theorists often include the results as independent
  measurements with a Gaussian distribution and disregard the boundary
  problem altogether.

@@

In \refsec{sample-based-det}, we assume that the PDF $P$ describe the underlying physics accurately,
and that the experiment observes each event with perfect accuracy. In order to estimate systematic
uncertainties, we lift these assumptions.

@@

\begin{figure}
    \includegraphics[width=.6\textwidth]{fig-topology.pdf}
    \caption{Decay topology for decays $B\to P_1 P_2 \ell_1 \bar\ell_2$. \label{fig:topology}}
\end{figure}

So far, we assumed that our underlying physical process was accurately described by our PDF $P$.
However, in several interesting processes we might only have an approximative result for $P$.
In this section we focus on one particular class of mismodelling of the signal PDF: the angular
momentum-cutoff in partial wave expansions. This mismodelling potentially affects a large
number of decays and scattering processes. For the sake of clarity we take the interesting class
of four-body decays $B\to P_1 P_2 \ell_1 \bar\ell_2$ as an example\footnote{This includes the rare $b\to s$ mediated $B$ decay $B \to K\pi\ell^+\ell^-$, and
the $V_{ub}$ suppressed decay $B\to \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in
the small-width approximation and when assuming a pure $P$-wave, resonant final state. An extension to
$S-P$ interference has been studied for $B\to K\pi\ell^+\ell^-$ \cite{Blake:2012mb,Becirevic:2011bp},
and for $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$ \cite{Faller:2013dwa}. For a first study of $S$, $P$
and $D$ interference, see \cite{Das:2014sra}}.\\

Within existing analyses, the PDFs for these decays are usually expressed in terms of one or a few
partial waves of the dimeson system. However, the angular momentum of the dimeson system is unbounded
from above, and gives rise to an infinite set of angular observables.\\

For the selected class of decays, we can describe that problem as follows: The PDF $P$ has a fixed dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokpill} for details
on the angular distribution.) However, on the level of decay amplitudes the dimeson system can have an arbitrarily
large total angular momentum $j$; only its third component is restricted to $m = -1,0,+1$.
We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(|m|)}$
with respect to the remaining angle $\theta_2$. Using the observables $S_k(\vec{\nu})$, $k=1,\dots,9$, with $\vec{\nu}=(q^2, k^2)$
and as defined in \refapp{partial-waves}, the expansion reads
\begin{equation}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,|m|}} S_{k,j}(\vec{\nu}) p_{j}^{(|m|)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where the the normalization factor $n_{j,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokpill})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,\\
        1\,, & k = 3,4,5,7,8,9\,,
    \end{cases}\,.
\end{equation}
The angular observables $S_{k,j}$ defined as coefficients of this expansion in Legendre polynomials
have the merit of a well-defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section. To wit,
ignoring further (orthogonal) terms in the PDF cannot affect the mean vector $(\dots, S_k, \dots)$ nor its variance.\\

Unfortunately, this benefit on the experimental side is accompanied with a theoretical draw back. Each observable
$S_{k,j}(\vec{\nu})$ consists of an infinite sum of bilinear of partial-wave amplitudes. It remains for theoretical
analyses to estimate or calculate the impact of partial waves beyond the S and P wave contributions.
(For $B\to K\pi\ell^+\ell^-$ a first study has been carried out where contributions up to the D wave are studied \cite{Das:2014sra}).\\

@@

The determination of a detector's angular acceptance is generally a difficult
task. Here, we will show a systematic method to estimate its effects
based on both the method of moments, and MC simulations of the
detector.\\

In the ideal case, one would have an explicit probabilistic
model of the detector acceptance effects and could thus write down
the full forward model from which the measured events arise. In practice,
this would require much more CPU time than available, and one is
therefore forced to simplify the model. The standard approach is to
generate \emph{true} particle events $\lbrace x_t^{(k)}\rbrace =
\lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ from a PDF
assumed to describe the bare physical process, and to propagate
those particles through a detailed simulation of the detector. The
observable traces that the particles leave in the detector are fed
into reconstruction algorithms resulting in the \emph{detector}
events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,
\vec\theta^{(k)}_d)\rbrace$.\\

As before, we assume the PDF $P$ can be decomposed into $N$ parts with angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$,
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{j = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
This assumption is based on a smooth angular acceptance function. While this will not be true for the laboratory frame,
smoothness may be recovered in a boosted frame, e.g. the rest frame of the decaying particle.
In that case, the matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For a fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace = \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ from the PDF $P_j$,
        \begin{equation}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector, in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Determine the uncorrected observables $\lbrace \est{S_k} \rbrace$ based on the detector
        events $\lbrace x^{(k)}_d\rbrace$. For all intents and purposes, automation of this step is only feasible
        with to the method of moments, since it does not suffer form convergence problems.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}

@@

Our finding can be summarized as follows:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the pull distribution of the observable.
        All pull distributions can successfully be fitted to a gaussian distribution. As an example, we
        show the pull distributions for the observables $S_5$ and $S_7$ for a SM-like $B\to K^*\ell^+\ell^-$
        decays in \reffig{pulls}.
    \item We study the absolute uncertainty $\sigma_k(K)$ with respect to the angular observable $S_k$
        as a function of the number of simulated events $K$. As expected for a multivariate gaussian
        distribution, we find that the absolute uncertainty is well fitted by
        \begin{equation}
            \label{eq:unc-on-mean}
            \sigma_k(K) = \frac{\sigma_k(1)}{\sqrt{K}}
        \end{equation}
        with $\sigma_k(1) = \order{1}$, regardless of the absolute size of $S_k$. The latter can best be shown
        for the uncertainties of two exampular observables. Taking again $S_5$ ($\simeq 27\%$) and $S_7$ ($\simeq 2\%$)
        for SM-like $B\to K^*\ell^+\ell^-$ decays, we show the absolute uncertainty in \reffig{errors}.
    \item For all toy studies, we produce plots of their skewness and their kurtosis as functions of
        $N$. We find in all cases that the mean skewness and kurtosis converge towards zero.
        \todo{Plot Skewness and Kurtosis for $S_5$ and $S_7$ as functions of the number of simulated events $K$.}
\end{itemize}
We find that all of the above results show consistently that the joint distrbution of the angular observables
converges towards a multivariate Gaussian distribution. We propose to describe the resulting distribution as
a multivariate Gauss when skewness and kurtosis are small enough. The precise targets for these quantities
are subject to the details of the physical analysis, and therefore outside the scope of our work.

@@

The PDF for the decay --- in the presence of Standard-Model operators and their chirality-flipped counter parts -- reads \cite{Boeer:2014xx}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma} /\, \rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \frac{3}{8\pi} \sum_i \frac{K_{i}(q^2)}{\rmdx{\Gamma} /\, \rmdx{q^2}} f_i(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{equation}
where $q^2$ denotes the dilepton mass squared, $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and
$N\pi$ systems, respectively, and $\theta_3 = \phi$ denotes the azimuthal angle.
The basis of angular functions reads \cite{Boeer:2014xx}
\begin{equation}
    \begin{aligned}
% 1
        f_{1ss} & = \sin^2\theta_1 &
        f_{1cc} & = \cos^2\theta_1 &
        f_{1c}  & = \cos\theta_1\\
% 2
        f_{2ss} & = \sin^2\theta_1 \cos\theta_2 &
        f_{2cc} & = \cos^2\theta_1 \cos\theta_2 &
        f_{2c}  & = \cos\theta_1   \cos\theta_2 \\
% 3
        f_{3sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\theta_3 &
        f_{3s}  & = \sin\theta_1              \sin\theta_2 \cos\theta_3 \\
% 4
        f_{4sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\theta_3 &
        f_{4s}  & = \sin\theta_1              \sin\theta_2 \sin\theta_3
    \end{aligned}
\end{equation}
The decay width reads
\begin{equation}
    \frac{\rmdx{\Gamma}} {\rmdx{q^2}} = 2 K_{1ss} + K_{1cc}
\end{equation}
In terms of the angles $\theta_1$ and $\theta_2$, instead of the cosines, the PDF reads
\begin{equation}
    P(q^2, \theta_1, \theta_2, \theta_3) = \frac{3}{8\pi} \sum_i S_i(q^2) f_i(\cos\theta_1, \cos\theta_2, \theta_3) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_i(q^2) \equiv K_i(q^2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.\\

>>>>

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[svgnames]{xcolor}
\usepackage{hepparticles}
\usepackage{hepnicenames}
\usepackage{hepunits}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}

@@

Integration over the non-angular variables yields
\begin{equation}
    \langle S_i\rangle
    \equiv \int \rmdx{^M \nu} S_i(\nuvec)
    = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N \theta} P(\nuvec,\thvec) \dual{f}_i(\thvec) \right].
\end{equation}
The remainder of this section describes the method of moments, in
which we replace the analytical integration by Monte Carlo (MC)
estimates.  The central tenet of MC integration is the fact that the
expectation value $E_P[g]$ of some function $g(x)$ under the
probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int \rmdx{x} P(x) g(x),
\end{equation}
can be approximated by the consistent and unbiased
estimator $\est{E_P[g]}$~\cite[sec. 8.2]{James:2006zz}
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1}^K g(x^{(k)}) \,,\,    x^{(k)} \sim P
\end{equation}
due to the strong law of large numbers for $K \to \infty$, assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$, are distributed as
$P$.
Throughout this letter we denote all MC estimators with a wide hat.\\

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain observables integrated over certain
bins of $\nuvec$. We define
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & \equiv \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} S_i(\nuvec)\\
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta} P(\nuvec,\thvec) \dual{f}_i(\thvec)  \right]\\
    & = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta}  P(\nuvec,\thvec) \dual{f}_i(\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \right],
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
For notational simplicity, let us forget about the $\nuvec$
integration and consider only $S_i$. In the limit $K \to \infty$, the
central limit theorem (CLT) implies that the random vector
\begin{equation}
  \label{eq:angular-obs-vec}
  \vecest{S} \equiv (S_1, \dots, S_i\,,
  \dots)
\end{equation}
follows a multivariate Gaussian distribution $\mathcal{N}( \vec{S},
\widehat{\Sigma})$ centered on the true value $\vec{S}$ with the
covariance $\widehat{\Sigma}_{ij}$ estimated as
\begin{equation}
    \cov[S_i,S_j] \to \est{\cov}[{S}_i, {S}_j]
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[\dual{f}_i(x^{(k)}) - \est{ S_i}\big]\,\big[\dual{f}_j(x^{(k)}) - \widehat{ S_j}\big]\,.
\end{equation}
In our physics applications, the parameter space is compact and each
$\dual{f}_i$ is bounded. Hence the requisites for the most basic
version of the CLT to hold --- finite mean and covariance of
$\dual{f}_i$ --- are automatically satisfied.  We would like to
emphasize that the sample covariance should \fred{``Should'' as in it
  does in practice? Reference toy studies below} rapidly converge
towards the true covariance matrix.  We propose to declare convergence
\fred{I'd prefer: check that skewness and kurtosis are sufficiently
  small} once estimators for skewness and kurtosis are sufficiently
small,

@@

  For a likelihood fit, the constraint needs to be part of the
  analysis performed by the experimental collaboration, and the
  resulting likelihood as a function of $\vec{S}$ is distinctly not
  Gaussian.  For experimentalists, communicating such a result has
  proved to be challenging due to technical reasons (data formats,
  size etc.) and the fear that theorists misinterprete the many
  nuisance parameters (from control channels, detector calibrations
  etc.) that are also part of the likelihood. This leads to the
  undesirable situation that that only the mode and standard errors
  are reported, and theorists often include the results as independent
  measurements with a Gaussian distribution and disregard the boundary
  problem as well as correlations altogether.

@@

In \refsec{sample-based-det}, we assume that the PDF $P$ describes the underlying physics accurately,
and that the experiment observes each event with perfect accuracy. In order to estimate systematic
uncertainties, we lift these assumptions.

@@

\begin{figure}
    \includegraphics[width=.6\textwidth]{fig-topology.pdf}
    \caption{Decay topology for decays $B\to P_1 P_2 \ell_1 \bar\ell_2$. \label{fig:topology}\fred{We should use the same symbol $\theta$ as in text}}
\end{figure}

In several interesting processes we might only have an approximate
result for $P$.  In this section, we focus on one particular class of
mismodelling of the signal PDF: the angular momentum-cutoff in partial
wave expansions. This mismodelling potentially affects a large number
of decays and scattering processes. For the sake of clarity we take
the interesting class of four-body decays $B\to P_1 P_2 \ell_1
\bar\ell_2$ \fred{How general is this? The plot implies that the
  dimeson and the dilepton do not have a common vertex. But this
  requires an intermediate state, e.g., $K^{*}$}as an
example\footnote{This includes the rare $b\to s$ mediated $B$ decay $B
  \to K\pi\ell^+\ell^-$, and the $V_{ub}$ suppressed decay $B\to
  \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in
  the small-width approximation and when assuming a pure $P$-wave,
  resonant final state. An extension to $S-P$ interference has been
  studied for $B\to K\pi\ell^+\ell^-$
  \cite{Blake:2012mb,Becirevic:2011bp}, and for $\bar{B}\to
  \pi\pi\ell^-\bar\nu_\ell$ \cite{Faller:2013dwa}. For a first study
  of $S$, $P$
  and $D$ interference, see \cite{Das:2014sra}}.\\

Within existing analyses, the PDFs of these decays are usually expressed in terms of one or a few
partial waves of the dimeson system. However, the angular momentum of the dimeson system is unbounded
from above, and gives rise to an infinite set of angular observables.\\

For the selected class of decays, we can describe that problem as follows: The PDF $P$ has a fixed dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokpill} for details
on the angular distribution.) However, at the level of decay amplitudes the dimeson system can have an arbitrarily
large total angular momentum $j$; only its third component is restricted to $m = -1,0,+1$. \fred{check}
We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(|m|)}$
with respect to the remaining angle $\theta_2$. Using the observables $S_k(\vec{\nu})$, $k=1,\dots,9$, with $\vec{\nu}=(q^2, k^2)$
and as defined in \refapp{partial-waves} \fred{A lot of mixing between appendices C and E}, the expansion reads
\begin{equation}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,|m|}} S_{k,j}(\vec{\nu}) p_{j}^{(|m|)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where the the normalization factor $n_{j,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokpill})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,\\
        1\,, & k = 3,4,5,7,8,9\,,
    \end{cases}\,.
\end{equation}
The angular observables $S_{k,j}$ defined as coefficients of this expansion in Legendre polynomials
have the merit of a well defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section. To wit,
ignoring further (orthogonal) terms in the PDF cannot affect the mean vector $(\dots, S_k, \dots)$ nor its variance. \fred{yes!}\\

Unfortunately, this benefit on the experimental side is accompanied with a theoretical draw back. Each observable
$S_{k,j}(\vec{\nu})$ consists of an infinite sum of bilinears of partial-wave amplitudes. It remains for theoretical
analyses to estimate or calculate the impact of partial waves beyond the S and P wave contributions.
(For $B\to K\pi\ell^+\ell^-$ a first study has been carried out where contributions up to the D wave are studied \cite{Das:2014sra}).\\

@@

The determination of a detector's angular acceptance is generally a difficult
task. Here, we will show a systematic method to estimate its effects
based on both the method of moments and MC simulations of the
detector.\\

In the ideal case, one would have an explicit probabilistic model of
the detector acceptance effects and could thus write down the full
forward model from which the measured events arise. In practice, this
would require much more CPU time than available, and one is therefore
forced to simplify the model. The standard approach is to generate
\emph{true} particle events $\lbrace x_t^{(k)}\rbrace =
\lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ from a PDF
assumed to describe the bare physical process, and to propagate those
particles through a detailed simulation of the detector. The
observable traces that the particles leave in the detector are fed
into reconstruction algorithms resulting in the \emph{detector} events
$\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,
\vec\theta^{(k)}_d)\rbrace$.\fred{What about uncertainty in $\vec{\nu}, \vec{\theta}$? How does it enter the experimental likelihood? And how our method?}\\

As before, we assume the PDF $P$ can be decomposed into $N$ parts with angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$, \fred{We don't need the $^{\text true}$.}
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{j = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
This assumption is based on a smooth angular acceptance function. While this will not be true for the laboratory frame,
smoothness may be recovered in a boosted frame, e.g. the rest frame of the decaying particle.
In that case, the matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For a fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace = \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ from the PDF $P_j$,
        \begin{equation}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector, in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Determine the uncorrected observables $\lbrace \est{S_k} \rbrace$ based on the detector
        events $\lbrace x^{(k)}_d\rbrace$. For all intents and purposes, automation of this step is only feasible
        with to the method of moments, since it does not suffer form convergence problems.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}

@@

Our finding can be summarized as follows:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the pull distribution of the observable.
        All pull distributions can successfully be fitted to a gaussian distribution. As an example, we
        show the pull distributions for the observables $S_5$ and $S_7$ for a SM-like $B\to K^*\ell^+\ell^-$
        decays in \reffig{pulls}.
    \item We study the absolute uncertainty $\sigma_k(K)$ with respect to the angular observable $S_k$
        as a function of the number of simulated events $K$. As expected for a multivariate gaussian
        distribution, we find that the absolute uncertainty is well fitted by
        \begin{equation}
            \label{eq:unc-on-mean}
            \sigma_k(K) = \frac{\sigma_k(1)}{\sqrt{K}}
        \end{equation}
        with $\sigma_k(1) = \order{1}$, regardless of the absolute size of $S_k$. The latter can best be shown
        for the example of uncertainties of two observables. Taking again $S_5$ ($\simeq 27\%$) and $S_7$ ($\simeq 2\%$)
        for SM-like $B\to K^*\ell^+\ell^-$ decays, we show the absolute uncertainty in \reffig{errors}.
    \item For all toy studies, we produce plots of their skewness and their kurtosis as functions of
        $N$. We find in all cases that the mean skewness and kurtosis converge towards zero.
        \todo{Plot Skewness and Kurtosis for $S_5$ and $S_7$ as functions of the number of simulated events $K$.}
\end{itemize}
We find that all of the above results show consistently that the joint distrbution of the angular observables
converges towards a multivariate Gaussian distribution. We propose to describe the resulting distribution as
a multivariate Gauss when skewness and kurtosis are small enough. The precise targets for these quantities
are subject to the details of the physical analysis, and therefore outside the scope of our work.

@@

The PDF for the decay --- in the presence of Standard-Model operators and their chirality-flipped counter parts --- reads \cite{Boeer:2014xx}
\begin{equation}
    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma} /\, \rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}} = \frac{3}{8\pi} \sum_i \frac{K_{i}(q^2)}{\rmdx{\Gamma} /\, \rmdx{q^2}} f_i(\cos\theta_1, \cos\theta_2, \theta_3)\,,
\end{equation}
where $q^2$ denotes the dilepton mass squared, $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and
$N\pi$ systems, respectively, and $\theta_3 = \phi$ denotes the azimuthal angle.
The basis of angular functions reads \cite{Boeer:2014xx}
\begin{equation}
    \begin{aligned}
% 1
        f_{1ss} & = \sin^2\theta_1 &
        f_{1cc} & = \cos^2\theta_1 &
        f_{1c}  & = \cos\theta_1\\
% 2
        f_{2ss} & = \sin^2\theta_1 \cos\theta_2 &
        f_{2cc} & = \cos^2\theta_1 \cos\theta_2 &
        f_{2c}  & = \cos\theta_1   \cos\theta_2 \\
% 3
        f_{3sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\theta_3 &
        f_{3s}  & = \sin\theta_1              \sin\theta_2 \cos\theta_3 \\
% 4
        f_{4sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\theta_3 &
        f_{4s}  & = \sin\theta_1              \sin\theta_2 \sin\theta_3
    \end{aligned}
\end{equation}
The decay width reads
\begin{equation}
    \frac{\rmdx{\Gamma}} {\rmdx{q^2}} = 2 K_{1ss} + K_{1cc}
\end{equation}
In terms of the angles $\theta_1$ and $\theta_2$, instead of the cosines, the PDF reads
\begin{equation}
    P(q^2, \theta_1, \theta_2, \theta_3) = \frac{3}{8\pi} \sum_i S_i(q^2) f_i(\cos\theta_1, \cos\theta_2, \theta_3) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_i(q^2) \equiv K_i(q^2) / (\rmdx{\Gamma} /\, \rmdx{q^2})$.\\
